Molecular dynamics simulation of synchronization in driven particles

### 2020-10-05 manuscript/manuscript_ajp_2020.tex
updated text in model/simulation

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  With
  numerical simulations, we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  We use molecular dynamics simulations to model the particle dynamics,
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We study 
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.

@@

%Outline
%describe types of physical systems 
Numerical simulations of confined, driven particles
can describe a diverse variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

Narrow channels studies are useful to provide insights 
of how particles move through geometries
such as wires, microchannels, and biological systems such as neuron axons and capillaries [more detail and references].
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Many such systems execute local oscillations
about stable points [elaborate].
%
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
The simulations use overdamped dynamics with Euler's method in a classical two-dimensional model for particles confined to a periodic potential. With only a few particles in two-dimensions, the simulation runs quickly, making this an appropriate model for undergraduates to explore.

Particles interact by the Yukawa potential.

A single particle has the equation of motion 
\begin{center}
$\eta \vec{v_i} = \vec{F}_{landscape,i} + \sum_{j}^{} \vec{F}_{ij} + \vec{F}_{Drive}(t)$.\\ 
\end{center}
This includes the force of the landscape, inter-particle interactions $\vec{F}_{ij} = - \nabla V(\vec{r}_{ij})$, and a driving force $\vec{F}_{Drive}(t) = [F_{DC} + F_{AC} \sin(\omega t)] \hat{y}$, where $\omega = 2 \pi f$. The landscape is a sine function that is position dependent and calculate the landscape force on the particle directly from its location. The driving force does add energy into the system, and some of it is lost. The damping comes from the viscous fluid providing a nonconservative force.

@@

%\section{MD Simulations}

\subsection{Single particle system}  % In subsection headings, only the first word is capitalized.

@@

\subsection{Kinked system}
\label{DispEqSection}	% You can label sections for reference

>>>>

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  With
  numerical simulations, we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  We use molecular dynamics simulations to model the particle dynamics,
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We study 
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

%Outline
%describe types of physical systems 
Numerical simulations of confined, driven particles
can be used to model a diverse variety of physical phenomena.
Particles which interact over long distances
include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more detail and references]
Particles which interact over short distances include
bubble arrays/emulsions [more systems and references].

Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and fluid microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

@@

\section{Molecular Dynamics Model}
\label{sec:MD}
We use a classical two-dimensional model for 
studying particle dynamics. 
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5$.
An individual particle has
position $\vec{r}_i = x_i \hat{i} + y_i \hat{j}$.
We use periodic boundary conditions
such that a particle leaving the edges of the system is mapped
back to a position within the simulation 
by the rules $x_i+L=x_i$ and $y_i+L=y_i$.

Particles are subject to external driving forces
$F_{Drive}(t)$
that change as a function of time.
We create several model light fields,
creating a landscape of potential minima and maxima
that modify the local force on a particle as a function of position
$\vec{F}_{landscape,i} = \vec{F}_{landscape}(\vec{r}_i)$
These landscape potential are static,
with fixed minima and maxima
that are periodic or quasi-periodic,
as described in Sec.~\ref{sec:results}.

We model particle interactions with
the Yukawa potential $\vec{F}_{ij} = \nabla V_{ij}(r_{ij})$,
\begin{equation}
V_{ij}(r_{ij}) = \frac{E_0}{r_{ij}} e^{-\kappa r_{ij}},
\end{equation}
where $r_{ij} = |\vec{r}_i - \vec{r}_j|$.
This is a
screened Coulomb potential
$E_0=2$ scales strength of repulsion
where $E_0 = k q_1 q_2$  [CHECK scaling/units].
$\kappa = 1/R_0$ is the screening parameter,
that screening accounts for the lengthscale at
which many particle
interactions and local environment
that reduces the interaction range of individual particles.
We fix $R_0$ to be unity in our simulation units.

Particle dynamics are
modeled with an overdamped
equation of motion
integrated with the Verlet method.
A single particle has the equation of motion 
\begin{equation}
\eta \vec{v_i} = \vec{F}_{landscape,i} + \sum_{i \neq j}^{N} \vec{F}_{ij} + \vec{F}_{Drive}(t).
\end{equation}
where $\eta = 1$.
Since $\vec{a}$ is zero,
the Verlet method simplifies to 
the Euler method
$\vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)$.

The driving force $\vec{F}_{Drive}(t) = [F_{DC} + F_{AC} \sin(\omega t)] \hat{y}$, where frequency is defined as $\omega = 2 \pi f$.
We modify $f$ and observe the change in dynamics.
The landscape is a sine function that is position dependent and calculate the landscape force on the particle directly from its location. The driving force does add energy into the system, and some of it is lost. The damping comes from the viscous fluid providing a nonconservative force.

@@

\section{Results}
\label{sec:results}

\subsection{Single particle system}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

@@

\subsection{Kinked system}
\label{sec:static}	% You can label sections for reference
